---
title: "NYC 311 SERVICE REQUEST"
author: "Nuno Fernandes"
date: "31/10/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import Libraries
```{r, echo=FALSE,warning=FALSE, message=FALSE, results=FALSE}
packages <- c("dlookr","Hmisc","funModeling","tidyverse","lubridate","ggplot2","hrbrthemes","xts","fpp2","TSstudio","forecast","shiny","shinyWidgets","shinythemes","plotly")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)
```

### read csv
```{r}
df<-read.csv("C:\\Users\\nunok\\Documents\\Msc Data Science\\data_viz\\project\\Customer_Service_Module.csv")
```

# Preprocessing

### convert char to Date and create duration var
```{r}
df$created_as_time <- parse_date_time(df$Created.Date, '%m/%d/%Y %I:%M:%S %p')
df$closed_as_time  <- parse_date_time(df$Closed.Date, '%m/%d/%Y %I:%M:%S %p')

df$duration_time   <- as.numeric(difftime(df$closed_as_time, df$created_as_time)) 
```

#### mean of duration to solve per day of the call
```{r}
#created_as_time convert to Date
df$Date <- as.Date(df$created_as_time)


df%>%
  group_by(df$Date)%>%
  mutate(mean_duration = mean(duration_time, na.rm = T))%>%
  distinct(Date, .keep_all = T) -> df_ts
```

### time series incident resolution time
```{r}
#time series of duration to solve per date
ts_a <- xts(x = df_ts$mean_duration, order.by = df_ts$Date)
```


### mean of duration until solved issue & number of calls per day 
```{r}
#if duration<0 => NA
df_ts2 <- df
df_ts2$duration_time[df_ts2$duration_time <0] <- NA


#created_as_time convert to Date
df_ts2$Date <- as.Date(df_ts2$created_as_time)

#compute mean duration & number of calls per day and filter
df_ts2%>%
  group_by(Date)%>%
  mutate(mean_duration = mean(duration_time, na.rm = T))%>%
  mutate(number = n())%>%
  distinct(Date, .keep_all = T) -> df_ts2

#to xts duration daily time series
ts_duration <- xts(x = df_ts2$mean_duration, order.by = df_ts2$Date)

#to xts number of call daily time series
ts_number <- xts(x = df_ts2$number, order.by = df_ts2$Date)
```

# Exploratory analysis

### convert xts to monthly ts time series 
```{r, echo=FALSE, warning=FALSE}
#convert xts duration to ts weekly time series
ts_monthly_duration <- to.monthly(ts_duration)
ts_monthly_duration <- xts_to_ts(ts_monthly_duration)

#convert xts number of call to ts time series
ts_monthly_number <- to.monthly(ts_number)
ts_monthly_number <- xts_to_ts(ts_monthly_number)

#plot
#autoplot(ts_monthly_duration)
#autoplot(ts_monthly_number)
```
### convert to weekly time series
```{r, echo=FALSE, warning=FALSE}
#convert xts duration to ts weekly time series
ts_weekly_duration <- to.weekly(ts_duration)
ts_weekly_duration <- xts_to_ts(ts_weekly_duration)

#convert xts number of call to ts time series
ts_weekly_number <- to.weekly(ts_number)
ts_weekly_number <- xts_to_ts(ts_weekly_number)

#plot
#autoplot(ts_weekly_duration)
#autoplot(ts_weekly_number)
```
### find weeks that seem to be an outliers
```{r}
#duration
#boxplot(ts_weekly_duration)
#tail(sort(ts_weekly_duration))
which(ts_weekly_duration==c(22030800, 29073600))

#number
#boxplot(ts_weekly_number)
#head(sort(ts_weekly_number))
which(ts_weekly_number==1)

```
### mean imputation(mean of previous and following week) & convert duration s to hr
```{r}
#duration
ts_weekly_duration[c(499,500)] <- mean(ts_weekly_duration[c(498,501)])
ts_weekly_duration[325] <- mean(ts_weekly_duration[c(324,326)])

#mean
ts_weekly_number[c(498,499, 500)] <- mean(ts_weekly_number[c(497,501)])

#duration from s to hours
ts_weekly_duration <- ts_weekly_duration/3600

```

### plot
```{r}
autoplot(cbind(ts_weekly_number,ts_weekly_duration))
```

### to percentage
```{r} 
#duration
ts_weekly_duration_percentage <- ((ts_weekly_duration - min(ts_weekly_duration))/(max(ts_weekly_duration)-min(ts_weekly_duration)))*100

#number
ts_weekly_number_percentage <- ((ts_weekly_number - min(ts_weekly_number))/(max(ts_weekly_number)-min(ts_weekly_number)))*100

```

### auto-arima duration
``` {r}
fit_duration <- auto.arima(ts_weekly_duration_percentage)
checkresiduals(fit_duration)
autoplot(fit_duration)
#autoplot(forecast(fit))
plot(fit_duration$x)
lines(fitted(fit_duration),col="blue")
```

### auto-arima number
``` {r}
# autoArima number
fit_number <- auto.arima(ts_weekly_number_percentage)
checkresiduals(fit_number)
autoplot(fit_number)
#autoplot(forecast(fit))
plot(fit_number$x)
lines(fitted(fit_number),col="blue")
```

### Bind the two plots
```{r}
autoplot(cbind(ts_weekly_number_percentage, ts_weekly_duration_percentage, fitted(fit_number), fitted(fit_duration) )) + ggtitle("Title") + labs(x = "Time", y= " Relative frequency (%)")
```

```{r}
number <- fitted(fit_number)
duration <-  fitted(fit_duration)

autoplot(cbind(number, duration)) + ggtitle("Weekly 311 calls") + labs(x = "Time", y= "Percentage of change") + scale_y_continuous(expand = c(0, 0), breaks = c(0,10,20,30,40,50,60,70)) + coord_cartesian(ylim=c(0, 70))+ theme(panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),panel.grid.major.x = element_blank() ) + theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                                               axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid')) +     scale_x_continuous(n.breaks = 10)
  
```

### ts to dataframe
```{r}
#ts to dataframe
df_duration <- fortify(ts_weekly_duration)
df_number <- fortify(ts_weekly_number)

#df to ts
ts_duration <- ts(df_duration)
ts_number <- ts(df_number)
```

### Why this decreasing trend in the duration since covid? Inspect number of cases/type and duration per case

### When was this sudden rise in calls? 
```{r}
#summary(df_ts2$number)
which(df_ts2$number>=3570)
df_ts2[2875,"Date"]
```

#seasonal plot
```{r}
ggmonthplot(ts_monthly_duration)+
  ylab("duration") +
  ggtitle("Seasonal plot: duration ")

ggmonthplot(ts_monthly_number)+
  ylab("number") +
  ggtitle("Seasonal plot: number ")

```

# weekly plot of n() Water Quality
```{r}
df_water = subset(df, Complaint.Type == "Water Quality")

df_water%>%
  group_by(Date)%>%
  mutate(number = n())%>% 
  distinct(Date, .keep_all = T) -> df_water_

ts_water_ <- xts(x = df_water_$number, order.by = df_water_$Date)

ts_weekly_water <- to.weekly(ts_water_)
ts_weekly_water <- xts_to_ts(ts_weekly_water)

autoplot(ts_weekly_water) +  ggtitle("Weekly 311 Water Quality Complaints") + labs(x = "Year", y= "Number") + scale_y_continuous(expand = c(0, 0), breaks = c(0,10,20,30,40,50)) + coord_cartesian(ylim=c(0, 60))+ theme(panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),panel.grid.major.x = element_blank() ) + theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                                               axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid')) +     scale_x_continuous(n.breaks = 10) + geom_line(colour = "darkgoldenrod")

```
# weekly plot of n() noise
```{r}
df_water = subset(df, Complaint.Type == "Noise")

df_water%>%
  group_by(Date)%>%
  mutate(number = n())%>% 
  distinct(Date, .keep_all = T) -> df_water_

ts_water_ <- xts(x = df_water_$number, order.by = df_water_$Date)

ts_weekly_water <- to.weekly(ts_water_)
ts_weekly_water <- xts_to_ts(ts_weekly_water)

autoplot(ts_weekly_water) +  ggtitle("Weekly 311 Noise complaints") + labs(x = "Year", y= "Number") + scale_y_continuous(expand = c(0, 0), breaks = c(0,50,100,150,200,250)) + coord_cartesian(ylim=c(0, 300))+ theme(panel.grid.minor.y = element_blank(),
panel.grid.minor.x = element_blank(),panel.grid.major.x = element_blank() ) + theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                                               axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid')) +     scale_x_continuous(n.breaks = 7) + geom_line(colour = "blueviolet")

```


# Create a data frame with one Borough
```{r, warning=FALSE}

#remove rows for unspecified Boroughs
c <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND")

df %>% filter(Borough %in% c) -> df

#df$Borough[df$Borough == "STATEN ISLAND"] <- "STATEN_ISLAND"

# BRONX first
df_bronx = subset(df, Borough == "BRONX")

df_ = subset(df_bronx, Complaint.Type == "Water System")

#compute mean duration & number of calls per day and filter
df_%>%
  group_by(Date)%>%
  mutate(number = n())%>% 
  distinct(Date, .keep_all = T) -> df_new

df_new$type = "Water System"
df_new$borough = "BRONX"

c = c("Noise", "Air Quality", "Lead", "Water Conservation", "Hazardous Materials", "Water Quality", "Sewer")

for (i in c){
  
  df_ = subset(df_bronx, Complaint.Type == i)
  #compute mean duration & number of calls per day and filter
  df_%>%
    group_by(Date)%>%
    mutate(number = n()) %>% 
    distinct(Date, .keep_all = T) -> df_

  df_$type = i
  df_$borough = "BRONX"
  
  #rbind df
  df_new = rbind(df_new, df_)
  
}
```

### Bind next Boroughs 
```{r}
list_borough <- c("BROOKLYN", "MANHATTAN", "STATEN ISLAND", "QUEENS")

list_complaint = c("Water System", "Noise", "Air Quality", "Lead", "Water Conservation", "Hazardous Materials", "Water Quality", "Sewer")

#outer-loop
for (borough in list_borough){
  df_ = subset(df, Borough == borough)
  #inner-loop
    for (complaint in list_complaint){
      df_2 = subset(df_, Complaint.Type == complaint)
        #compute mean duration & number of calls per day and filter
        df_2%>%
          group_by(Date)%>%
          mutate(number = n()) %>% 
          distinct(Date, .keep_all = T) -> df_2
        
        df_2$type = complaint
        df_2$borough = borough
        
        #rbind df
        df_new = rbind(df_new, df_2)
      
    }
}

```


# Plotly
```{r}

       ggplotly(df_new %>% filter(borough == "BROOKLYN") %>% 
                     ggplot(aes(Date, number, colour = Complaint.Type)) +  geom_line(size = 1, alpha = .9) +
                       #uncomment line below to show area under curves
                       #geom_area(aes(fill = symbol),position="identity",alpha=.2) +
                           theme_minimal(base_size=16) +
                           theme(axis.title = element_blank(),
                           plot.background  = element_rect(fill = "black"),
                           panel.background = element_rect(fill = "black"),
                           panel.grid       = element_blank(),
                           legend.text      = element_text(colour = "white")) 
                ) 
  

```

